Structural Divergence Analysis

### 2017-01-08 analysis-structure-core.Rmd
cambios 08/01/2017

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data.

@@

@@

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

@@

```{r_comparison_mean_z.RSD_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles
ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset)) +
  geom_line() +
  geom_text(data = subset(d, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8)+
  geom_vline(xintercept = active.sites, show.legend = F) +
  geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) + 
  scale_colour_manual(name = NULL, values = myColors)

@@

@@

knitr::kable(cor.df, digits = 2, caption = "Correlation between <Z-scores> profiles")
```

### Distances to the active sites

```{r_comparison_da}

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.ca, xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM, xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.ca)
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM)

cor.df = data.frame("comparison" = c("(exp - mut) vs min.da.CA", "(exp - mut) vs min.da.CM"), "R" = c(r.min.da.CA, r.min.da.CM))

knitr::kable(cor.df, digits = 2, caption = "Correlation between differences in <Z-scores> profiles and the minimal distance to the active site")
```

```{r_comparison_vs_plots}

>>>>

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data. 
For the core we considered sites with no gaps in the whole multiple alignment.

@@

```{r read_the_core, comment = ""}
# indexes.core
indexes.core = as.numeric(read.csv(file.path("DATA", paste("indexes.core", "_", family, ".csv", sep = "")))[1, -1])
```

@@

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_point() +
    geom_errorbar(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

@@

```{r-comparison-mean-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}

@@

# eliminate duplicates
active.sites = active.sites[!duplicated(active.sites)]
contacts.active.sites = contacts.active.sites[!duplicated(active.sites)]

# new.indexes
new.active.index = c()
for (i in (1:(length(active.sites)))){
  new.active.index[i] = which(indexes.core == active.sites[i])
}

new.contacts.active.index = c()
for (i in (1:(length(contacts.active.sites)))){
  new.contacts.active.index[i] = which(indexes.core == contacts.active.sites[i])
}

active.sites = new.active.index
contacts.active.sites = new.contacts.active.index

@@

knitr::kable(cor.df, digits = 2, caption = "Correlation between <Z-scores> profiles")

### 2017-01-08 analysis-structure-normal-modes.Rmd
cambios 08/01/2017

---
title: "Structural Divergence Analysis"
author: "María Laura Marcos, Julian Echave"
date: "25 de julio de 2016"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = 2, autodep = T)
```

```{r load_libraries}
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

setwd("C:/Users/Usuario/Desktop/VariabilidadEstructuralProteica/OUT")

source("out_subset_CA_ANM/multiplot.R")
source("out_subset_CA_ANM/my-functions.R")
```

```{r set_my_colors, eval = T}
myColors = c("blue", "darkgreen", "red", "red")
names(myColors) <- c("exp", "mut", "strong", "mut+sel")

```
# Introduction
We study the evolutionary divergence of proteins using the LFENM and compare with experimental data.

```{r set_run_variables}
# Set options
family = "globins"
R0 = 10
enm = "ANM"

# significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
data.dir = paste("out_subset_CA_ANM", sep = "")
```

```{r set_file_names}
## Input and pre-processing of data
# Set up input filenames

reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(family, "_out_m.da.csv", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

dri2.exp.fname = file.path(data.dir, paste(family, "_R0_10_beta_strong.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
Pn.exp.fname = file.path(data.dir, paste(family,"_R0_10_beta_strong.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
energy.exp.fname = file.path(data.dir, paste(family, "_R0_10_beta_strong.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))

dri2.strong.fname = file.path(data.dir, paste(family, "_R0_10_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
Pn.strong.fname = file.path(data.dir, paste(family, "_R0_10_beta_strong.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
energy.strong.fname = file.path(data.dir, paste(family, "_R0_10_beta_strong.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

dri2.mut.fname = file.path(data.dir,paste(family, "_R0_10_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
Pn.mut.fname = file.path(data.dir, paste(family, "_R0_10_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_10_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

@@

```{r set_reference_protein}
## which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
```

We use the `r enm` Elastic Network Model with `R0 =` `r R0` for the `r family` family. 
We use as reference for the ENM the protein with pdb code `r reference`.
The `r family` family consists of the following proteins:

```{r set_experimental_protein_names}
## Experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

[//]: Prepare site.info data frame

```{r set_up_site_info, eval = T}

## read site information of reference protein
site.info = read.csv(site.info.fname, header = T, sep = ";")

## add info to site.info
d = ddply(site.info, "site", .fun = function(x) data.frame("da" = min(x[, c(-1, -2)])))

## Define shells of active site neighborhood (first shell is from 2.5 to 7.5, second from 7.5 to 12.5, etc)
d$shell = cut(d$da, breaks = c(-.1, 2.5, 7.5, 12.5, 17.5, 22.5, 27.5, 32.5, 37.5, 42.5, 47.5, 52.5, max(d$da)))
levels(d$shell) = seq(length(levels(d$shell))) - 1 # rename shell levels
site.info = merge(site.info, d, by = "site")

## Read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

## Add pdb.info to site.info
site.info = cbind(site.info, pdb.info)
```

[//]: Prepare experimental data

```{r read_experimental_data}
# read experimental data
dri2.exp = read.csv(dri2.exp.fname, header = T)
site = site.info$site
dri2.exp = dri2.exp[, site]
rownames(dri2.exp) = protein.exp

Pn.exp = read.csv(Pn.exp.fname, header = T)
rownames(Pn.exp) = protein.exp

energy.exp = read.csv(energy.exp.fname, header = T)
rownames(energy.exp) = protein.exp

[//]: Prepare data for no selection (mut)

```{r read_and_set_data_mut}
# read simulated data

dri2.mut = read.csv(dri2.mut.fname, header = T)
site = site.info$site
dri2.mut = dri2.mut[, site]
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")

Pn.mut = read.csv(Pn.mut.fname, header = T)
rownames(Pn.mut) = paste("mut", rownames(Pn.mut), sep = "")

energy.mut = read.csv(energy.mut.fname, header = T)
rownames(energy.mut) = paste("mut", rownames(energy.mut), sep = "")
```

[//]: Prepare data for strong selection

```{r read_and_set_data_strong}
# read simulated data

dri2.strong = read.csv(dri2.strong.fname, header = T)
site = site.info$site
dri2.strong = dri2.strong[, site]
rownames(dri2.strong) = paste("strong", rownames(dri2.strong), sep = "")

Pn.strong = read.csv(Pn.strong.fname, header = T)
rownames(Pn.strong) = paste("strong", rownames(Pn.strong), sep = "")

energy.strong = read.csv(energy.strong.fname, header = T)
rownames(energy.strong) = paste("strong", rownames(energy.strong), sep = "")
```

```{r calculate MSD}
MSD.exp = mean(rowMeans(as.matrix(dri2.exp), na.rm = T), na.rm = T) 
MSD.mut = mean(rowMeans(as.matrix(dri2.mut)))
MSD.strong = mean(rowMeans(as.matrix(dri2.strong)))

df.MSD = data.frame(MSD.exp = MSD.exp, MSD.mut = MSD.mut, MSD.strong = MSD.strong)
knitr::kable(df.MSD, digits = 4)

# Structural divergence along sites

Here we study how structural divergence varies from site to site. We will compare simulated data and empirical data.

```{r prepare-site-data}
# Prepare my data.frame
# rows of dri2 are proteins, columns are sites

tdri2.strong = as.data.frame(t(as.matrix(dri2.strong)))
tdri2.strong = cbind(site.info, tdri2.strong)
siteComparison.strong = melt(tdri2.strong, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

siteComparison.all.datasets = rbind(
  data.frame("dataset" = "exp", siteComparison.exp),
  data.frame("dataset" = "mut", siteComparison.mut),
  data.frame("dataset" = "strong", siteComparison.strong))

@@

### Mean z.RSD vs site

To measure structural divergence we use z.RSD profiles, which are the best (most similar distributions and profiles w.r.t experimental).

The following plot compares the mean z.RSD profiles, including standard errors (with .999 confidence level for the mean)
*Warning: the effective number of data should be smaller than the actual data, especially for the experimental dataset because of several factors, including smoothing of the profiles, phylogenetic relationships, etc. We need to deal with this for quantitative analysis. Probably we'll need to use bootstrapping to calculate standard errors non-parametrically*

```{r mean-z-RSD-vs-site}
# Comparison of z.RSD profiles annotated with functional information (reported active sites).
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

# dat.with.means and standard errors
d = ddply(dat,c("dataset","site","AA","shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})

##

z.RSD.exp = d$mean.z.RSD[d$dataset == "exp"]
z.RSD.strong = d$mean.z.RSD[d$dataset == "mut+sel"]
z.RSD.mut = d$mean.z.RSD[d$dataset == "mut"]

m.min.da.CM.ca = read.csv(file.path(data.dir, "min.da.CM.ca.csv", sep = ""))[, 2:4]

data.csv = data.frame(m.min.da.CM.ca[1:151, ], "mean.z.RSD.exp" = z.RSD.exp, "mean.z.RSD.strong" = z.RSD.strong, "mean.z.RSD.mut" = z.RSD.mut)

write.csv(data.csv, file.path(data.dir, "data.da.csv", sep = ""))

m.min.da.CM = m.min.da.CM.ca$min.da.CM[1:151]

data.plot = data.frame("d.exp.mut" = z.RSD.mut - z.RSD.exp, "d.exp.strong" = z.RSD.mut - z.RSD.strong, "m.min.da.CM" = m.min.da.CM)

r = round(cor(data.plot$d.exp.mut, data.plot$m.min.da.CM), 2)

ggplot(data.plot, aes(y = d.exp.mut, x = m.min.da.CM)) +
  geom_point()+
  geom_smooth(method = "lm") +
  geom_label(aes(x = 5, y = 2, label = paste("R =", as.character(r)))) +
    labs(y = "mean.z.RSD.mut - mean.z.RSD.exp", x = "min.da.CM")

r = round(cor(data.plot$d.exp.strong, data.plot$m.min.da.CM), 2)

ggplot(data.plot, aes(y = d.exp.strong, x = m.min.da.CM)) +
  geom_point()+
  geom_smooth(method = "lm") +
  geom_label(aes(x = 5, y = 0.5, label = paste("R =", as.character(r)))) +
    labs(y = "mean.z.RSD.strong - mean.z.RSD.exp", x = "min.da.CM")

##

m.min.da.ca = m.min.da.CM.ca$min.da.ca[1:151]

data.plot = data.frame("d.exp.mut" = z.RSD.mut - z.RSD.exp, "d.exp.strong" = z.RSD.strong - z.RSD.exp, "m.min.da.ca" = m.min.da.ca)

r = round(cor(data.plot$d.exp.mut, data.plot$m.min.da.ca), 2)

ggplot(data.plot, aes(y = d.exp.mut, x = m.min.da.ca)) +
  geom_point()+
  geom_smooth(method = "lm") +
  geom_label(aes(x = 5, y = 2, label = paste("R =", as.character(r)))) +
    labs(y = "mean.z.RSD.mut - mean.z.RSD.exp", x = "min.da.ca")

r = round(cor(data.plot$d.exp.strong, data.plot$m.min.da.ca), 2)

ggplot(data.plot, aes(y = d.exp.strong, x = m.min.da.ca)) +
  geom_point()+
  geom_smooth(method = "lm") +
  geom_label(aes(x = 5, y = 2, label = paste("R =", as.character(r)))) +
    labs(y = "mean.z.RSD.strong - mean.z.RSD.exp", x = "min.da.ca")

#d$dataset = factor(d$dataset, levels = c("exp", "mut", "strong"))
d$label = as.character(d$site)
active.sites = d$site[d$shell == 0]

p = ggplot(d,aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name=NULL, values = myColors)

# Plot separate and comparative plots
p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")

p.facet %+% subset(d, dataset == "exp") 
p.facet %+% subset(d, dataset == "mut")
p.facet %+% subset(d, dataset == "mut+sel") 
p.facet
#p + theme(legend.position = "bottom", legend.title = element_blank())
p + theme(legend.position = "bottom")

VAL107 and its neighbors (including site 32) are more structurally conserved than predicted.

### Correlation between site profiles

```{r}
dat = siteComparison.all.datasets

# dat.with.means and standard errors
dat = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})

dat.exp = subset(dat, dataset == "exp")
dat.mut = subset(dat, dataset == "mut")
dat.strong = subset(dat, dataset == "strong")

ggplot(dat, aes(x = site, y = mean.z.RSD, col = dataset)) + geom_line() +
    geom_text(data = subset(dat, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 2)

d = dat.exp
d$d.exp.mut = dat.exp$mean.z.RSD - dat.mut$mean.z.RSD
d$d.exp.strong = dat.exp$mean.z.RSD - dat.strong$mean.z.RSD
d$d.strong.mut = dat.strong$mean.z.RSD - dat.mut$mean.z.RSD
dd = melt(data = d[, c("site", "se.z.RSD", "d.exp.mut", "d.strong.mut")], c("site", "se.z.RSD"))

ggplot(dd, aes(x = site, y = value, col = variable)) +
    geom_line() 
r = round(cor(d$d.strong.mut, d$d.exp.mut), 2)

ggplot(d, aes(x = scale(d.strong.mut), y = scale(d.exp.mut))) + 
    geom_point(aes(col = shell == 0)) +
    geom_smooth(method = "lm") +
    geom_label(aes(x = 0, y = 3, label = paste("R =", as.character(r)))) +
    labs(x = "predicted selection vs mutation (RSD change)",y="empirical vs mutation (RSD change)")

r.exp.mut = cor(dat.exp$mean.z.RSD, dat.mut$mean.z.RSD)
r.exp.strong = cor(dat.exp$mean.z.RSD, dat.strong$mean.z.RSD)
r.mut.strong = cor(dat.mut$mean.z.RSD, dat.strong$mean.z.RSD)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))
knitr::kable(cor.df, digits = 2)
write.csv(dat, "structure-comparison-profiles.csv")

# Structural divergence along normal modes
In this section we consider structural divergence using normal mode coordinates.
Pn is the square divergence along a normal mode relative to the mean.

[//]: Prepare mode data

```{r prepare-mode-data}
mode = seq(ncol(Pn.strong))

@@

modeComparison.strong = merge(energy.strong.long, Pn.strong.long, by = c("mode", "protein"))

modeComparison.strong = ddply(modeComparison.strong, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

@@

modeComparison.exp = merge(energy.exp.long,Pn.exp.long,by=c("mode", "protein"))

# renormalize Pn
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.strong <- na.omit(modeComparison.strong)
modeComparison.mut <- na.omit(modeComparison.mut)

[//]: Add quantile regression data

```{r}
## Quantile regression

@@

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "strong"))
```

### Pn vs. mode energy

@@

Here we show the square deviation along each normal mode (normalized so that <Pn> = 1) vs. the normal-mode's energy. For the empirical data, we use the actual energy obtained from Keff for the given comparison. It is this energy the one that should be used, since it's inverse is the expected value of Pn. Lines are the .5 (median), .05, and .95 quantiles obtained using `rq()`
- Upper panel: simulated vs empirical data, shown separately.
- Lower panel: simulated vs empirical overlaped and zoomed to better see the pattern.

```{r Pn-vs-mode-energy,cache = 2}
# Plot Pn vs energy
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

```{r median-Pn-vs-mode-energy, cache = 2}
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

### Correlation between mode profiles

```{r}
dat = modeComparison.all.datasets
#dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

[//]: Write output for further analysis with other programs

```{r write-data-to-output}
write.csv(siteComparison.exp, "site-structure-comparison-exp.csv",row.names=F)
write.csv(modeComparison.exp, "mode-structure-comparison-exp.csv",row.names=F)
```

>>>>

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

# Introduction
We study the evolutionary divergence of proteins using the LF-ENM and comparing simulated mutants with experimental data.

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

# load functions
source(file.path(data.dir, "multiplot.R"))
source(file.path(data.dir, "my-functions.R")) 
```

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

## profiles to compare
### Pn
Pn.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
Pn.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
Pn.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))

### energy
energy.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))
energy.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

@@

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r prepare_site_data}
# prepare the data
Pn.exp = read.csv(Pn.exp.fname, header = T)
rownames(Pn.exp) = protein.exp
energy.exp = read.csv(energy.exp.fname, header = T)
rownames(energy.exp) = protein.exp

Pn.mut = read.csv(Pn.mut.fname, header = T)
rownames(Pn.mut) = paste("mut", rownames(Pn.mut), sep = "")
energy.mut = read.csv(energy.mut.fname, header = T)
rownames(energy.mut) = paste("mut", rownames(energy.mut), sep = "")

Pn.strong = read.csv(Pn.strong.fname, header = T)
rownames(Pn.strong) = paste("strong", rownames(Pn.strong), sep = "")
energy.strong = read.csv(energy.strong.fname, header = T)
rownames(energy.strong) = paste("strong", rownames(energy.strong), sep = "")

## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## mut
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## strong
tdri2.strong = as.data.frame(t(as.matrix(dri2.strong)))
tdri2.strong = cbind(site.info, tdri2.strong)
siteComparison.strong = melt(tdri2.strong, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "strong", siteComparison.strong))

@@

### Structural divergence along normal modes

In this section we consider structural divergence using normal mode coordinates. Pn is the square divergence along a normal mode relative to the mean.

```{r-prepare-mode-data}
mode = seq(ncol(Pn.strong))

@@

modeComparison.strong = merge(energy.strong.long, Pn.strong.long, by = c("mode", "protein"))
modeComparison.strong = ddply(modeComparison.strong, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

@@

modeComparison.exp = merge(energy.exp.long,Pn.exp.long,by=c("mode", "protein"))
# renormalize Pn
modeComparison.exp = ddply(modeComparison.exp, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.strong <- na.omit(modeComparison.strong)
modeComparison.mut <- na.omit(modeComparison.mut)
```

```{r-add-quantile-regression-data}

## Quantile regression

@@

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "strong"))
```
### Pn vs. mode energy

@@

Here we show the square deviation along each normal mode (normalized so that = 1) vs. the normal-mode's energy. For the empirical data, we use the actual energy obtained from Keff for the given comparison. It is this energy the one that should be used, since it's inverse is the expected value of Pn. Lines are the .5 (median), .05, and .95 quantiles obtained using rq()

Upper panel: simulated vs empirical data, shown separately.
Lower panel: simulated vs empirical overlaped and zoomed to better see the pattern.

```{r-Pn-vs-energy}

# Plot Pn vs energy
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

```{r-median-vs-mode-energy}

dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

### correlaction between mode profiles

```{r-correlation-mode-profiles}
dat = modeComparison.all.datasets
#dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = dat.exp$Pn.median, 
     y = dat.mut$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "mut.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.mut$Pn.median)), max(c(dat.exp$Pn.median, dat.mut$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.mut$Pn.median)), max(c(dat.exp$Pn.median, dat.mut$Pn.median))))
abline(0, 1)

plot(x = dat.exp$Pn.median, 
     y = dat.strong$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "strong.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.strong$Pn.median)), max(c(dat.exp$Pn.median, dat.strong$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.strong$Pn.median)), max(c(dat.exp$Pn.median, dat.strong$Pn.median))))
abline(0, 1)

plot(x = dat.mut$Pn.median, 
     y = dat.strong$Pn.median, 
     xlab = "mut.Pn.median", 
     ylab = "strong.Pn.median", 
     xlim = c(min(c(dat.mut$Pn.median, dat.strong$Pn.median)), max(c(dat.mut$Pn.median, dat.strong$Pn.median))),
     ylim = c(min(c(dat.mut$Pn.median, dat.strong$Pn.median)), max(c(dat.mut$Pn.median, dat.strong$Pn.median))))
abline(0, 1)
```

### 2017-01-08 analysis-structure-window-contacts.Rmd
cambios 08/01/2017

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
```

>>>>

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.windows.contacts.rot.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.contacts.rot.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.contacts.rot.csv", sep = ""))
```
